---
title: "Statistical Performance Indicators - Analysis"
author: "Brian Stacy"
date: "5/8/2020"
output:
  html_document: 
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning = FALSE, message = FALSE)

# Load Packages
library(tidyverse)
library(skimr)
library(kableExtra)
library(readxl)
library(Hmisc)
library(haven)
library(gt)
library(xlsx)

# Directory for SPI csv files that were previously created
csv_dir <- "C:/Users/wb469649/Documents/Github/SPI_AKI/R/01_data/013_outputs"
out_dir <- "C:/Users/wb469649/Documents/Github/SPI_AKI/R/02_analysis/023_outputs"

```

# Introduction

This file produces a number of summary statistics for the SPI dimensions.  Summary statistics are presented both population weighted and unweighted.  We will produce a table of summary statistics with mean, median, standard deviation, etc, and also a table of means by region and income level.


```{r load, include=FALSE}

#import data
load(paste(csv_dir, "SPI.RData", sep="/" ))

SPI <- SPI %>%
  filter((lending %in% c('IDA', 'IBRD', 'Blend')) | (iso3c=='PSE')) #keep just the borrowing countries & West Bank & Gaza

```



# Summary Statistics

Below we show summary statistics for the year 2019.  The same table is available for other years in the appendix.  The first table shows the summary statistics weighted by population.  The second shows unweighted.

## Population Weighted Summary Statistics

```{r sumstats_weighted, echo=FALSE}

#add function to produce weighted summary stats
        my_skim<-    skim_with( numeric = sfl( mean = ~ wtd.mean(.,  weights =pop, na.rm=TRUE),
                                               sd = ~ sqrt(wtd.var(.,  weights=pop, na.rm=TRUE)),
                                               p25 = ~ (wtd.quantile(., probs=c(0.25),  weights=pop, na.rm=TRUE)),
                                               p50 = ~ (wtd.quantile(., probs=c(0.5), weights=pop, na.rm=TRUE)),
                                               p75 = ~ (wtd.quantile(., probs=c(0.75), weights=pop, na.rm=TRUE)),
                                               N = ~ sum(!is.na(.))))

for (reference_year in 2016:2019) {
temp <- SPI %>%
  filter(date==reference_year) 

#add population weight
pop <- temp$population

temp <- temp %>%
  select( SPI.OVRL.SCR, contains('SPI.D1'),contains('SPI.D2'),contains('SPI.D3'),contains('SPI.D4')) %>%
  my_skim() %>%
  yank('numeric') %>%
  rename(Series=skim_variable) %>%
  left_join(metadata) %>% #add labels to variables
  select(Series, 'Indicator Name', mean, sd,p0,p100, hist, N)

assign(paste("spi_weighted_sumstats",reference_year,sep="_"), temp)

}

#save these to an excel file for easy sharing
write.xlsx(
  spi_weighted_sumstats_2016,
  file=paste(out_dir, 'summary_statistics_SPI.xlsx', sep="/"),
  sheetName='spi_weighted_sumstats_2016',
  append=FALSE
) 

write.xlsx(
  spi_weighted_sumstats_2017,
  file=paste(out_dir, 'summary_statistics_SPI.xlsx', sep="/"),
  sheetName='spi_weighted_sumstats_2017',
  append=T
)     

write.xlsx(
  spi_weighted_sumstats_2018,
  file=paste(out_dir, 'summary_statistics_SPI.xlsx', sep="/"),
  sheetName='spi_weighted_sumstats_2018',
  append=T
)     

write.xlsx(
  spi_weighted_sumstats_2019,
  file=paste(out_dir, 'summary_statistics_SPI.xlsx', sep="/"),
  sheetName='spi_weighted_sumstats_2019',
  append=T
)     



#create grouping variable for table
group <- spi_weighted_sumstats_2019 %>%
  mutate(group=case_when(
    grepl('OVRL', Series) ~ "Overall",
    grepl('D1', Series) ~ "Dimension 1",
    grepl('D2', Series) ~ "Dimension 2",
    grepl('D3', Series) ~ "Dimension 3",
    grepl('D4', Series) ~ "Dimension 4"
  ))

kable(spi_weighted_sumstats_2019,
      digits=2,
      col.names=c('Series', 'Description', 'Mean','SD','Min','Max','Histogram', 'N'),
      caption="Statistical Performance Indicator Population Weighted Summary Statistics for 2019") %>%
  kable_styling(bootstrap_options = c('striped','hover')) %>%
  pack_rows(index=table(fct_inorder(group$group)))
      
      



```

## Regional Means (Weighted by Population)

```{r regional_weighted, echo=FALSE}

for (reference_year in 2016:2019) {
df_overall <- SPI %>%
  filter(date==reference_year) 
#produce by region
        sumstats<- df_overall %>%
            group_by(region) %>%
            filter(!is.na(region)) %>%
            select(region, SPI.OVRL.SCR, contains('SPI.D1'),contains('SPI.D2'),contains('SPI.D3'),contains('SPI.D4'), population) %>%
            summarise_all(~round(100*wtd.mean(.,population, na.rm=T),2)) %>%
            select(-population) %>%
            mutate(SPI.OVRL.SCR=round(SPI.OVRL.SCR/100,2),
                   SPI.D1.MSC=round(SPI.D1.MSC/100,2),
                   SPI.D2.CS=round(SPI.D2.CS/100,2),
                   SPI.D3.AKI=round(SPI.D3.AKI/100,2),
                   SPI.D4.DPO=round(SPI.D4.DPO/100,2)) 
        
        #produce global number
        sumstats_gl<- df_overall %>%
            mutate(region='Global') %>%
            group_by(region) %>%
            select(region, SPI.OVRL.SCR, contains('SPI.D1'),contains('SPI.D2'),contains('SPI.D3'),contains('SPI.D4'), population) %>%
            summarise_all(~round(100*wtd.mean(.,population, na.rm=T),2)) %>%
            select(-population) %>%
            mutate(SPI.OVRL.SCR=round(SPI.OVRL.SCR/100,2),
                   SPI.D1.MSC=round(SPI.D1.MSC/100,2),
                   SPI.D2.CS=round(SPI.D2.CS/100,2),
                   SPI.D3.AKI=round(SPI.D3.AKI/100,2),
                   SPI.D4.DPO=round(SPI.D4.DPO/100,2))
        
        
        #transpose data
        sumstats_df_long <-sumstats_gl %>%
            bind_rows(sumstats) 
        
        sumstats_df <- as.data.frame(t(sumstats_df_long %>% select(-region)))
        colnames(sumstats_df) = sumstats_df_long$region 
        
        
        sumstats_df <- sumstats_df %>%
            rownames_to_column() %>%
            rename(series=rowname)
        
        
        #create labels df
        metadata_tab2_overall <- metadata %>% 
            janitor::clean_names() %>%
            select(series, indicator_name)
        
        
        #add variable label
        sumstats_df <- sumstats_df %>%
            left_join(metadata_tab2_overall) %>%
            rename(Series=series,
                   Label=indicator_name) %>%
            select(Series, Label, everything())

 assign(paste("spi_weighted_regional",reference_year,sep="_"), sumstats_df)
       
        }
 #save these to an excel file for easy sharing
write.xlsx(
  spi_weighted_regional_2016,
  file=paste(out_dir, 'summary_statistics_SPI.xlsx', sep="/"),
  sheetName='spi_weighted_regional_2016',
  append=T
) 

write.xlsx(
  spi_weighted_regional_2017,
  file=paste(out_dir, 'summary_statistics_SPI.xlsx', sep="/"),
  sheetName='spi_weighted_regional_2017',
  append=T
) 
write.xlsx(
  spi_weighted_regional_2018,
  file=paste(out_dir, 'summary_statistics_SPI.xlsx', sep="/"),
  sheetName='spi_weighted_regional_2018',
  append=T
) 
write.xlsx(
  spi_weighted_regional_2019,
  file=paste(out_dir, 'summary_statistics_SPI.xlsx', sep="/"),
  sheetName='spi_weighted_regional_2019',
  append=T
) 
                
        DT::datatable(spi_weighted_regional_2019, caption=htmltools::tags$caption(
            style = 'caption-side: top; text-align: left;',
            htmltools::em("Population Weighted Mean of SPI Indicators by Region in 2019.  Sub-indicators scaled to be 0-100."))    ,
            extensions = 'Buttons', options=list(
                dom = 'Bfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                pageLength = 60)) 

```


## Unweighted Summary Statistics

```{r sumstats_unweighted, echo=FALSE}

        my_skim<-    skim_with( numeric = sfl( mean = ~ mean(.,   na.rm=TRUE),
                                               sd = ~ sqrt(var(.,   na.rm=TRUE)),
                                               p25 = ~ (quantile(., probs=c(0.25),   na.rm=TRUE)),
                                               p50 = ~ (quantile(., probs=c(0.5),  na.rm=TRUE)),
                                               p75 = ~ (quantile(., probs=c(0.75),  na.rm=TRUE)),
                                               N = ~ sum(!is.na(.))))

for (reference_year in 2016:2019) {
temp <- SPI %>%
  filter(date==reference_year) %>%
  select( SPI.OVRL.SCR, contains('SPI.D1'),contains('SPI.D2'),contains('SPI.D3'),contains('SPI.D4')) %>%
  my_skim() %>%
  yank('numeric') %>%
  rename(Series=skim_variable) %>%
  left_join(metadata) %>% #add labels to variables
  select(Series, 'Indicator Name', mean, sd,p0,p100, hist, N)

assign(paste("spi_unweighted_sumstats",reference_year,sep="_"), temp)

}

#save these to an excel file for easy sharing
write.xlsx(
  spi_unweighted_sumstats_2016,
  file=paste(out_dir, 'summary_statistics_SPI.xlsx', sep="/"),
  sheetName='spi_unweighted_sumstats_2016',
  append=T
) 

write.xlsx(
  spi_unweighted_sumstats_2017,
  file=paste(out_dir, 'summary_statistics_SPI.xlsx', sep="/"),
  sheetName='spi_unweighted_sumstats_2017',
  append=T
)      

write.xlsx(
  spi_unweighted_sumstats_2018,
  file=paste(out_dir, 'summary_statistics_SPI.xlsx', sep="/"),
  sheetName='spi_unweighted_sumstats_2018',
  append=T
) 
write.xlsx(
  spi_unweighted_sumstats_2019,
  file=paste(out_dir, 'summary_statistics_SPI.xlsx', sep="/"),
  sheetName='spi_unweighted_sumstats_2019',
  append=T
) 

#create grouping variable for table
group <- spi_unweighted_sumstats_2019 %>%
  mutate(group=case_when(
    grepl('OVRL', Series) ~ "Overall",
    grepl('D1', Series) ~ "Dimension 1",
    grepl('D2', Series) ~ "Dimension 2",
    grepl('D3', Series) ~ "Dimension 3",
    grepl('D4', Series) ~ "Dimension 4"
  ))

kable(spi_unweighted_sumstats_2019,
      digits=2,
      col.names=c('Series', 'Description', 'Mean','SD','Min','Max','Histogram', 'N'),
      caption="Statistical Performance Indicator Unweighted Summary Statistics for 2019") %>%
  kable_styling(bootstrap_options = c('striped','hover')) %>%
  pack_rows(index=table(fct_inorder(group$group)))
      
      



```

## Regional Means (Unweighted)

```{r regional_unweighted, echo=FALSE}

for (reference_year in 2016:2019) {
df_overall <- SPI %>%
  filter(date==reference_year) 
#produce by region
        sumstats<- df_overall %>%
            group_by(region) %>%
            filter(!is.na(region)) %>%
            select(region, SPI.OVRL.SCR, contains('SPI.D1'),contains('SPI.D2'),contains('SPI.D3'),contains('SPI.D4'), population) %>%
            summarise_all(~round(100*mean(., na.rm=T),2)) %>%
            select(-population) %>%
            mutate(SPI.OVRL.SCR=round(SPI.OVRL.SCR/100,2),
                   SPI.D1.MSC=round(SPI.D1.MSC/100,2),
                   SPI.D2.CS=round(SPI.D2.CS/100,2),
                   SPI.D3.AKI=round(SPI.D3.AKI/100,2),
                   SPI.D4.DPO=round(SPI.D4.DPO/100,2)) 
        
        #produce global number
        sumstats_gl<- df_overall %>%
            mutate(region='Global') %>%
            group_by(region) %>%
            select(region, SPI.OVRL.SCR, contains('SPI.D1'),contains('SPI.D2'),contains('SPI.D3'),contains('SPI.D4'), population) %>%
            summarise_all(~round(100*mean(., na.rm=T),2)) %>%
            select(-population) %>%
            mutate(SPI.OVRL.SCR=round(SPI.OVRL.SCR/100,2),
                   SPI.D1.MSC=round(SPI.D1.MSC/100,2),
                   SPI.D2.CS=round(SPI.D2.CS/100,2),
                   SPI.D3.AKI=round(SPI.D3.AKI/100,2),
                   SPI.D4.DPO=round(SPI.D4.DPO/100,2))
        
        
        #transpose data
        sumstats_df_long <-sumstats_gl %>%
            bind_rows(sumstats) 
        
        sumstats_df <- as.data.frame(t(sumstats_df_long %>% select(-region)))
        colnames(sumstats_df) = sumstats_df_long$region 
        
        
        sumstats_df <- sumstats_df %>%
            rownames_to_column() %>%
            rename(series=rowname)
        
        
        #create labels df
        metadata_tab2_overall <- metadata %>% 
            janitor::clean_names() %>%
            select(series, indicator_name)
        
        
        #add variable label
        sumstats_df <- sumstats_df %>%
            left_join(metadata_tab2_overall) %>%
            rename(Series=series,
                   Label=indicator_name) %>%
            select(Series, Label, everything())

 assign(paste("spi_unweighted_regional",reference_year,sep="_"), sumstats_df)
       
        }
 #save these to an excel file for easy sharing
write.xlsx(
  spi_unweighted_regional_2016,
  file=paste(out_dir, 'summary_statistics_SPI.xlsx', sep="/"),
  sheetName='spi_unweighted_regional_2016',
  append=T
) 

write.xlsx(
  spi_unweighted_regional_2017,
  file=paste(out_dir, 'summary_statistics_SPI.xlsx', sep="/"),
  sheetName='spi_unweighted_regional_2017',
  append=T
)     

write.xlsx(
  spi_unweighted_regional_2018,
  file=paste(out_dir, 'summary_statistics_SPI.xlsx', sep="/"),
  sheetName='spi_unweighted_regional_2018',
  append=T
)     
write.xlsx(
  spi_unweighted_regional_2019,
  file=paste(out_dir, 'summary_statistics_SPI.xlsx', sep="/"),
  sheetName='spi_unweighted_regional_2019',
  append=T
)     
DT::datatable(spi_unweighted_regional_2019, caption=htmltools::tags$caption(
            style = 'caption-side: top; text-align: left;',
            htmltools::em("Unweighted Mean of SPI Indicators by Region in 2019.  Sub-indicators scaled to be 0-100."))    ,
            extensions = 'Buttons', options=list(
                dom = 'Bfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                pageLength = 60)) 

```

# Appendix

## Summary Statistics 2016

Below we show summary statistics for the year 2016.

```{r sumstats_2016, echo=FALSE}

kable(spi_weighted_sumstats_2016,
      digits=2,
      col.names=c('Series', 'Description', 'Mean','SD','Min','Max','Histogram', 'N'),
      caption="Statistical Performance Indicator Population Weighted Summary Statistics for 2016") %>%
  kable_styling(bootstrap_options = c('striped','hover')) %>%
  pack_rows(index=table(fct_inorder(group$group)))

DT::datatable(spi_weighted_regional_2016, caption=htmltools::tags$caption(
            style = 'caption-side: top; text-align: left;',
            htmltools::em("Population Weighted Mean of SPI Indicators by Region in 2016.  Sub-indicators scaled to be 0-100."))    ,
            extensions = 'Buttons', options=list(
                dom = 'Bfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                pageLength = 60)) 

kable(spi_unweighted_sumstats_2016,
      digits=2,
      col.names=c('Series', 'Description', 'Mean','SD','Min','Max','Histogram', 'N'),
      caption="Statistical Performance Indicator Unweighted Summary Statistics for 2016") %>%
  kable_styling(bootstrap_options = c('striped','hover')) %>%
  pack_rows(index=table(fct_inorder(group$group)))


        
DT::datatable(spi_unweighted_regional_2016, caption=htmltools::tags$caption(
            style = 'caption-side: top; text-align: left;',
            htmltools::em("Unweighted Mean of SPI Indicators by Region in 2016.  Sub-indicators scaled to be 0-100."))    ,
            extensions = 'Buttons', options=list(
                dom = 'Bfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                pageLength = 60))         

```



## Summary Statistics 2017

Below we show summary statistics for the year 2017.

```{r sumstats_2017, echo=FALSE}
kable(spi_weighted_sumstats_2017,
      digits=2,
      col.names=c('Series', 'Description', 'Mean','SD','Min','Max','Histogram', 'N'),
      caption="Statistical Performance Indicator Population Weighted Summary Statistics for 2017") %>%
  kable_styling(bootstrap_options = c('striped','hover')) %>%
  pack_rows(index=table(fct_inorder(group$group)))

DT::datatable(spi_weighted_regional_2017, caption=htmltools::tags$caption(
            style = 'caption-side: top; text-align: left;',
            htmltools::em("Population Weighted Mean of SPI Indicators by Region in 2017.  Sub-indicators scaled to be 0-100."))    ,
            extensions = 'Buttons', options=list(
                dom = 'Bfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                pageLength = 60)) 

kable(spi_unweighted_sumstats_2017,
      digits=2,
      col.names=c('Series', 'Description', 'Mean','SD','Min','Max','Histogram', 'N'),
      caption="Statistical Performance Indicator Unweighted Summary Statistics for 2017") %>%
  kable_styling(bootstrap_options = c('striped','hover')) %>%
  pack_rows(index=table(fct_inorder(group$group)))

DT::datatable(spi_unweighted_regional_2017, caption=htmltools::tags$caption(
            style = 'caption-side: top; text-align: left;',
            htmltools::em("Unweighted Mean of SPI Indicators by Region in 2017.  Sub-indicators scaled to be 0-100."))    ,
            extensions = 'Buttons', options=list(
                dom = 'Bfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                pageLength = 60))               
```



## Summary Statistics 2018

Below we show summary statistics for the year 2018.

```{r sumstats_2018, echo=FALSE}
kable(spi_weighted_sumstats_2018,
      digits=2,
      col.names=c('Series', 'Description', 'Mean','SD','Min','Max','Histogram', 'N'),
      caption="Statistical Performance Indicator Population Weighted Summary Statistics for 2018") %>%
  kable_styling(bootstrap_options = c('striped','hover')) %>%
  pack_rows(index=table(fct_inorder(group$group)))

DT::datatable(spi_weighted_regional_2018, caption=htmltools::tags$caption(
            style = 'caption-side: top; text-align: left;',
            htmltools::em("Population Weighted Mean of SPI Indicators by Region in 2018.  Sub-indicators scaled to be 0-100."))    ,
            extensions = 'Buttons', options=list(
                dom = 'Bfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                pageLength = 60)) 

kable(spi_unweighted_sumstats_2018,
      digits=2,
      col.names=c('Series', 'Description', 'Mean','SD','Min','Max','Histogram', 'N'),
      caption="Statistical Performance Indicator Unweighted Summary Statistics for 2018") %>%
  kable_styling(bootstrap_options = c('striped','hover')) %>%
  pack_rows(index=table(fct_inorder(group$group)))

DT::datatable(spi_unweighted_regional_2018, caption=htmltools::tags$caption(
            style = 'caption-side: top; text-align: left;',
            htmltools::em("Unweighted Mean of SPI Indicators by Region in 2018.  Sub-indicators scaled to be 0-100."))    ,
            extensions = 'Buttons', options=list(
                dom = 'Bfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                pageLength = 60))               
```
