---
title: "Statistical Performance Index - Availability of Key Indicators"
output:
  html_document: default
  toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(wbstats)
library(httr)
library(jsonlite)
library(leaflet)
library(ggcorrplot)
library(skimr)
library(rsdmx)

work_dir<-"C:/Users/wb469649/Documents/Github/SPI_AKI/Documentation"

```

# Overview

DECIS has been tasked with coming up with a set of roughly 12 indicators for the Availability of Key Indicators (AKI) dimension of the Statistical Performance Index (SPI).  Based on feedback from the concept note of the SPI, the SPI team has requested the indicators for the AKI:  

  1.	Cover topic areas (i.e. the SDGs) as well as possible (for instance, in the current set of indicators below there is no coverage of environmental SDGs).    Tier 1 SDG indicators are preferred  
  2.	The indicators must be (as much as possible) produced by the country statistical systems, so as to measure statistical capacity.  Not modelled estimates.
  3.	Included in the WDI.  


I am continually updating a dashboard to help us with the indicator selection here:
https://datanalytics.worldbank.org/content/435

## Statistical Performance Index Overview

The SPI framework is designed to capture different aspects of national statistical capacity by employing most relevant and representative variables that are publicly available. The SPI can be used to gauge statistical performance of individual countries over time or cross-country comparisons of performance at a point in time. 

The SPI aims to provide an objective, justifiable/verifiable assessment of the statistical performance of countries over time by using publicly available information from international agencies and country websites that were produced by national statistical systems. The SPI framework helps countries and development partners identify the strengths and weaknesses of national statistical systems and areas of potential improvements. It could also provide actionable guidance for national statistical systems in areas that may require further and deeper assessment. 
Key characteristics of the SPI are:  

  - Uses only publicly accessible data  
  - Transparent methodology  
  - Easily replicable  
  - Provides a long-time series to track progress in performance  
  - Captures outcomes and supporting elements   
  - Reflects the SDGs.   
  - Facilitates at-a-glance comparisons on a global scale  
  
A recent paper by Cameron et al. (2019) provides the conceptual foundation for SPI in measuring the ability of NSS to produce high quality data to inform national and international policy decisions  


### General Notes

We could consider adding a metric on whether administrative data or private sector data is utilized by the statistical system.  This is what a modern statistical system should be able to do.

## Availability of Key Indicators (AKI) Overview

Transforming source data into statistical outputs (indicators) and releasing them on a timely basis shows that the statistical systems are utilizing their capacity in data production. Reporting relevant data to specialized international agencies on time and getting them published in their respective databases demonstrates that statistical systems meet required quality standards and timeliness. Therefore, this dimension evaluates national statistical systems by reviewing the availability of country data for the most recent year in international databases. By looking at the data availability in international databases, it also makes the assessment cost-effective.

The selection of the indicators under the AKI dimension is based on the following principles:  

  - It should provide some assessment of national statistical performance, i.e. the indicators should be produced by countries;  
  - The availability of indicators should be verifiable, with established standards and methodologies in producing the statistics.  
  - It should help address the development concerns of countries, especially with SDGs. Nine indicators included in this category overlap to large extent with the Tier 1 SDG list that are conceptually clear, established methodology and standards are available and data regularly produced by countries;  
  
  
These selected indicators cover key socio-economic and SDG indicators that have well established standards and methodology in the area of poverty, health, education, and economic development. It is assumed that if national statistical systems have capacity in the first two dimensions they should be able to produce these selected indicators. 


# Initial Set of Indicators

The initial proposal for the AKI is below.  The way the AKI index is scored is that countries are given a value of '1' if the indicator is available for the country and '0' otherwise.  The total AKI score for the country is then the 

```{r aki_orig, echo=FALSE, message=FALSE, warning=FALSE}

aki_orig<-read_csv(file=paste(work_dir, "aki_indicator_original.csv", sep="/"))

kable(aki_orig,
      caption = "Initial Proposed Set of AKI indicators by SPI Team",
      col.names = c('Indicator', 'Score 1 - Data is available once in three years, prior to reference year', 'Score 0'),
      format='markdown') 



```






```{r data_pulls, include=FALSE}
#########
# Figure out the Source ID for SDGs
#########
# Get all the source names from the World Bank API
sourceRequest <- GET(url = "http://api.worldbank.org/v2/sources?per_page=500&format=json")
sourceResponse <- content(sourceRequest, as = "text", encoding = "UTF-8")

# Parse the JSON content and convert it to a data frame.
sourceJSON <- fromJSON(sourceResponse, flatten = TRUE) %>%
  data.frame()

# Print the data frame and review the source names and ids
cols <- c("id","name")
sourceJSON[,cols]


##########
# Pull Tags for SDG Data
##########

# make request to World Bank API
indicatorRequest <- GET(url = "http://api.worldbank.org/v2/indicator?per_page=500&format=json&source=46")
indicatorResponse <- content(indicatorRequest, as = "text", encoding = "UTF-8")

# Parse the JSON content and convert it to a data frame.
indicatorsJSON <- fromJSON(indicatorResponse, flatten = TRUE) %>%
  data.frame()

# Print and review the indicator names and ids from the dataframe
cols <- c("id","name")
#indicatorsJSON[,cols] # to view all the indicator names, remove the # at the beginning of this line. Please note that there are over 500 indicators.


##########
# Pull Tags for WDI Data
##########

# make request to World Bank API
wdiRequest <- GET(url = "http://api.worldbank.org/v2/indicator?per_page=20000&format=json&source=2")
wdiResponse <- content(wdiRequest, as = "text", encoding = "UTF-8")

# Parse the JSON content and convert it to a data frame.
wdisJSON <- fromJSON(wdiResponse, flatten = TRUE) %>%
  data.frame()

# Print and review the indicator names and ids from the dataframe
cols <- c("id","name")
#indicatorsJSON[,cols] # to view all the indicator names, remove the # at the beginning of this line. Please note that there are over 500 indicators.


##########
# Pull Tags for EdStats Data
# Use different syntax than before
##########

# make request to UIS API
EdStatsRequest <- GET(url = "http://api.worldbank.org/v2/indicator?per_page=20000&format=json&source=12")
EdStatsResponse <- content(EdStatsRequest, as = "text", encoding = "UTF-8")

# Parse the JSON content and convert it to a data frame.
EdStatsJSON <- fromJSON(EdStatsResponse, flatten = TRUE) %>%
  data.frame()

# Print and review the indicator names and ids from the dataframe
cols <- c("id","name")
#indicatorsJSON[,cols] # to view all the indicator names, remove the # at the beginning of this line. Please note that there are over 500 indicators.

##########
# Pull for UIS UNESCO
##########


# need to input your API key here
#uis_key <- rstudioapi::askForSecret("UIS_key")

#now make request using API
url = paste("https://api.uis.unesco.org/sdmx/data/SDG4/STU.PT.L1+L2._T._T+F+M._T.G2_3+GLAST.INST_T._Z.READING+MATH._Z._Z._T._T._T._Z._Z._T.?startPeriod=2010&endPeriod=2019&format=sdmx-compact-2.1&subscription-key=",rstudioapi::askForSecret("UIS_key"), sep="")
uisRequest <- GET(url)

#build database
uis_df <- readSDMX(url, isURL=T) %>%
  as.data.frame()

# do some minor data cleaning to build official indicator name (i don't know why this isn't automatically in the API pulls!)
uis_df <- uis_df %>%
  mutate(
    subj=case_when(
      SUBJECT=="READING" ~ "Read",
      SUBJECT=="MATH" ~ "Math",
      TRUE ~ as.character(NA)
    ),
    grd_level=case_when(
      GRADE=="G2_3" & EDU_LEVEL=="L1" ~ "G2t3",
      GRADE=="GLAST" & EDU_LEVEL=="L1" ~ "Primary",
      GRADE=="GLAST" & EDU_LEVEL=="L2" ~ "LowerSec",
      TRUE ~ as.character(NA)
    )
  ) %>%
  mutate(INDICATOR_ID=paste(subj, grd_level, sep=".")) %>%
  mutate(INDICATOR_ID=case_when(
    SEX=="_T" ~ INDICATOR_ID,
    SEX=="F" ~ paste(INDICATOR_ID,"F", sep="."),
    SEX=="M" ~ paste(INDICATOR_ID,"M", sep="."),
    TRUE ~ as.character(NA)
  ))


#pull world bank country info to merge
country_info <- wbcountries()

uis_df<-uis_df %>%
  mutate(iso2c=REF_AREA) %>%
  left_join(country_info) %>%
  select(colnames(country_info), TIME_PERIOD, OBS_VALUE)

```








# Final Set of Indicators After Analysis

After consultation with WDI team, we decided to compose the AKI from 15 indicators chosen from the first 10 SDGs, 2 from environmental indicators, and 3 from national accounts.  

## New AKI Indicator Set

Choosing among the indicators we received from WDI experts on important indicators, below is the chosen list:  

  1. **No Poverty:** Poverty headcount ratio at $1.90 a day (2011 PPP) (% of population)  
  2. **Zero Hunger:**  Prevalence of undernourishment (% of population)
  3. **Good Health and Well-Being:** Mortality rate, under-5 (per 1,000 live births)	 
  4. **Quality Education:**  Primary completion rate, total (% of relevant age group)
  5. **Gender Equity:**  Number of maternal deaths	
  6. **Clean Water and Sanitation:**  People using safely managed drinking water services (% of population)	
  7. **Affordable and Clean Energy:**  Access to electricity (% of population)
  8. **Decent Work and Economic Growth:**  GDP per capita growth (annual %)	
  9. **Industry, Innovation, and Infrastructure:**  Research and development expenditure (% of GDP)	
  10. **Reduced Inequalities:**  Annualized average growth rate in per capita real survey mean consumption or income, bottom 40% of population (%)
  11. **Environment:**  Level of water stress: freshwater withdrawal as a proportion of available freshwater resources
  12. **Environment:**  Renewable electricity output (% of total electricity output)
  13. **National Accounts:**  Debt service (PPG and IMF only, % of exports of goods, services and primary income)	
  14. **National Accounts:**  	Consumer price index (2010 = 100)
  15. **National Accounts:**  Manufacturing, value added (% of GDP)


## Notes  

  - Poverty headcount ratio at national poverty lines (% of population) was changed to Poverty headcount ratio at $1.90 a day (2011 PPP) (% of population) because only 20% of countries had the former indicator, including wealthy countries such as the United States.  Roughly 40% have the latter indicator  
  - 76% of countries have an undernourishment indicator, so this indicator can still distinguish countries
  - Maternal mortality was considered, but these data tend to be quite old, even for wealthy countries. Only 3% of countries had data within 3 years of 2018.  Mortality rate under 5 was chosen.
  - Either primary or secondary completion were reasonable candidates.  Given the Education GPs focus on learning poverty at primary level, primary completion could be something that is slightly more relevant over the 5-10 year horizon  
  - The suggested gender indicator was the gender parity index for secondary school enrollment.  Given that school completion was used in the education indicator, I switched to the female share of employment in senious and middle management  
  - Originally we had basic drinking water as an indicator, but almost all countries already collect this data, so it had little power to distinguish countries.  I switched to People using safely managed drinking water services (% of population) which has more variation across countries.  
  - Access to electricity was recommended for affordable and clean energy by WDI team.  
  - GDP per capita growth (annual %)	was chosen for decent work and economic growth.  
  - Research and development expenditure (% of GDP)	was chosen for industry, innovation, and infrastructure at recommendation of WDI experts  
  - Annualized average growth rate in per capita real survey mean consumption or income, bottom 40% of population (%) was recommended by WDI experts  
  - Level of water stress: freshwater withdrawal as a proportion of available freshwater resources was recommended by WDI experts.  
  - Renewable electricity output (% of total electricity output) was recommended by WDI experts  
  - Debt service (PPG and IMF only, % of exports of goods, services and primary income)	recommended by WDI experts  
  - Manufacturing, value added (% of GDP) original AKI indicator chosen by SPI team. Another option was the GDP price deflator, but this indicator would be perfectly correlated with GDP per capita growth which is already included.
  

### Affordable and Clean Energy

Access to electricity is the easiest.  These are typically based on a household survey (DHS, MICS).  Could theoretically do a breakdown for census versus survey data.  Pretty much every government, every year collects this data, but the quality is low in many cases.  In cases of low quality, this data is not used.  

Some surveys (DHS, MICS) only ask a binary question of whether or not electricity available.  Some others offer richer data on quality, reliability, etc.  We could consider factoring these into our index.

### Clean Water and Sanitation Notes 

Safely managed drinking water has less coverage than basic drinking water.  The hurdle for countries is measuring water quality (test for E coli, etc).  Infrastructrure tends to be more static than other indicators, so expanding the window for the frequency of data may be needed.  May consider indicator system where we have 0 points for no data, 0.5 points for all data except water quality (this is available in a pdf report by the Bank and Unicef), and 1 point for full indicator for safely managed water.

### Gender Equity

Top tier indicators are:  
  - Maternal Mortality  (Vital statistics are foundational to a statistical system)
  - Gender Violence  (This is something imporant for statistical systems to build)
  
Second Tier:  
  - Age of Marriage  
  - Sexual reproductive
  - Time use of females
  
### Education  

Top tier indicator would be 4.1.1.  Could be built using HLO.  JP recommended a 4 year window.  Something like 1 point for having within a 4 year window.  0.5 points for within an 8 year window, 0 otherwise

Recommneded possibility of covering spectum of SDG 4 indicators and creating an index across all SDGs.  THis would include all Tier 1 and Tier 2 indicators in an index.
  
### Economics Statistics

Pull from Proposal: Macroeconomic Dashboard to measure progress towards SDG Goal 17/Target 17.13

* External Sector  
* Financial Sector  
* Fiscal Sector  
* Real Sector and Prices   
* Commondity Prices (Export/Import)  



## Updated Scoring Methodology

  - The scoring methodology was altered in this iteration.  
  - We have switched from scoring '1' if in a three year window of 2018 and '0'.  
  - The new scoring methodology includes frequency as a criteria to help distinguish between country statistical systems  

The score for an individual indicator $j$ for a country $i$ is now the fraction of times in a five year window an indicator is available.

$$ AKI_i^j=\frac{\sum_{t}^{5}{A_{it}^j}}{5} $$

For instance, if the reference window is 2018 and say a country has GDP data available in 2013, 2015, and 2017, then the country will get 0.6 points on this indicator.  


```{r updated_AKI }
aki<-c('Poverty headcount ratio at $1.90 a day (2011 PPP) (% of population)',
        'Prevalence of undernourishment (% of population)',
        'Mortality rate, under-5 (per 1,000 live births)',
        'Primary completion rate, total (% of relevant age group)',
        'Number of maternal deaths',
        'People using safely managed drinking water services (% of population)',
        'Access to electricity (% of population)',
       'GDP per capita growth (annual %)',
       'Research and development expenditure (% of GDP)',
       'Annualized average growth rate in per capita real survey mean consumption or income, bottom 40% of population (%)',
       'Level of water stress: freshwater withdrawal as a proportion of available freshwater resources',
       'Renewable electricity output (% of total electricity output)',
       'Debt service (PPG and IMF only, % of exports of goods, services and primary income)',
       'Consumer price index (2010 = 100)',
       'Manufacturing, value added (% of GDP)'
)

get_tag_aki_df<-wdisJSON %>%
  filter((name %in% aki	)) %>%
  arrange(factor(name, levels = aki)) %>%
  select(id, name,  sourceOrganization)


aki_list<-get_tag_aki_df[,'id']

reference_year=2018
df_aki <-wb(country="countries_only", 
            indicator=aki_list,
            startdate=reference_year-4,
            enddate=reference_year,
            return_wide = T) %>%
        mutate_at(.vars=aki_list, ~if_else(is.na(.),0,1)) %>% #create 0,1 variable for whether data point exists for country
        group_by(country) %>%
        summarise_all((~(if(is.numeric(.)) sum(., na.rm = TRUE)/5 else first(.)))) %>% #group by country to create one observation per country containing whether or not data point existed
        # Availability of Key Indicator Country Score=Weighted Score÷Maximum Category Score ×100
        mutate(AKI=100*rowMeans(.[aki_list], na.rm=T)) %>%
        mutate(ISO_A3=iso3c) 

```

## Metadata on Updated AKI Indicators

```{r metadata, echo=FALSE}

      kable(get_tag_aki_df,
      caption = "Metadata of Proposed Set of AKI indicators",
      format='markdown') 

```



## Summary Statistics of Updated AKI Indicators

- Below we show the summary statistics of the AKI indicators  
- These indicators have been converted based on the new scoring methodology which includes frequency as a criteria to help distinguish between country statistical systems.  Indicator values range from 0 to 1. 

```{r sumstats, echo=FALSE}
        #add function to produce weighted summary stats
        my_skim<-    skim_with( numeric = sfl( mean = ~ mean(.,   na.rm=TRUE),
                                               sd = ~ sqrt(var(.,   na.rm=TRUE)),
                                               p25 = ~ (quantile(., probs=c(0.25),   na.rm=TRUE)),
                                               p50 = ~ (quantile(., probs=c(0.5),  na.rm=TRUE)),
                                               p75 = ~ (quantile(., probs=c(0.75),  na.rm=TRUE)),
                                               complete = ~ sum(!is.na(.)))) 
        sumstats <- df_aki %>%
            select(aki_list) 
        
        #produce summary stats
        sumstats_df<-my_skim(sumstats) %>%
            yank("numeric") %>%
            mutate(id=skim_variable) %>%
            select(id, mean, sd, p0, p25, p50, p75, p100, complete,  hist) 
        
        #create labels df
        get_tag_aki_labs<-wdisJSON %>%
            filter(name %in% aki) %>%
            arrange(factor(name, levels = aki)) %>%
            select(id, name)
        
        
        #add variable label
        sumstats_df <- sumstats_df %>%
            left_join(get_tag_aki_labs) %>%
            select(id, name,  mean, sd, p0, p25, p50, p75, p100, complete, hist)
        
      kable(sumstats_df,
      caption = "Summary Statistics Across Countries of Proposed Set of AKI indicators by WDI Team",
      digits=2,
      format='markdown') 
        
        
```

## Correlations of Updated AKI Indicators

```{r correlations, echo=FALSE, fig.height=16, fig.width=18}

names(aki_list)=aki

#calculate correlations between teacher practices
df_corr_plot <-    round(cor(select(df_aki,aki_list), use="complete.obs"), 2)

#plot the correlation in a nicely formatted table
pcorr<- ggcorrplot(df_corr_plot,
                   outline.color = "white",
                   ggtheme = theme_bw(),
                   colors = c("#F8696B", "#FFEB84", "#63BE7B"),
                   legend.title = "Correlation",
                   title = "Correlation Between AKI Indicators",
                   lab=T) + 
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 20)) +
  theme_bw() +
  theme(
    text = element_text(size = 12),
  )

pcorr


```

# Appendix 




# Full Set of Indicators Recommended by WDI team

Additionally, led by Hiroko, we have a list of SDG indicators here suggested by the WDI team based on the criteria above.
Below are the recommendations by the WDI team.

```{r aki_rec, echo=FALSE, message=FALSE, warning=FALSE}

aki_rec<-read_csv(file=paste(work_dir, "aki_indicator_recommendations.csv", sep="/")) 
  

kable(aki_rec,
      caption = "Proposed Set of AKI indicators by WDI Team",
      format='markdown') 



```
